---
title: "Lauryna_plots"
editor: visual
execute: 
  echo: false
  warning: false
  error: false
format: 
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    fig-width: 6
    fig-height: 4
---

```{r}
library(dplyr)
library(openxlsx)
library(readxl)
library(ggsurvfit)
library(tidycmprsk)
library(survival)
library(survminer)
library(autoReg)
library(ggstatsplot)
library(tidyr)
library(DescTools)
library(forcats)
library(ggpubr)
library(rstatix)
library(agricolae)
library(EnvStats)
library(ggsci)
```

```{r}
theme_set(theme_bw()+
            theme(panel.grid = element_blank()))

HNF1B_simple <- read_excel("C:/Coding/studentai/Lauryna/HNF1B_curated Nr.2.xlsx")

HNF1B_simple$diagnosis <- 1

HNF1B_simple$age_pre <- as.numeric(HNF1B_simple$age_pre)
HNF1B_simple$age_post <- as.numeric(HNF1B_simple$age_post)

HNF1B_simple <- HNF1B_simple %>%
  mutate(mutacijos_tipas = ifelse(grepl('17q12', HNF1B.mutation..localization..nucleotide.and.amino.acid.change., ignore.case = TRUE), 'cdeletion', mutacijos_tipas))

```

```{r}
HNF1B_simple<-HNF1B_simple %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == '3b (Schwartz)', '3.0', CKD.stage...7)) %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == 'IIIa', '3.0', CKD.stage...7)) %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == 'IIIb', '3.0', CKD.stage...7)) %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == '3a', '3.0', CKD.stage...7)) %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == '3b', '3.0', CKD.stage...7)) %>%
  mutate(CKD.stage...7 = ifelse(CKD.stage...7 == '0.0', NA, CKD.stage...7)) %>%
  mutate(CKD.stage...7 = as.numeric(CKD.stage...7))

HNF1B_simple <- HNF1B_simple %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == '3b (Schwartz)', '3.0', CKD.stage...20)) %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == 'IIIa', '3.0', CKD.stage...20)) %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == 'IIIb', '3.0', CKD.stage...20)) %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == '3a', '3.0', CKD.stage...20)) %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == '3b', '3.0', CKD.stage...20)) %>%
  mutate(CKD.stage...20 = ifelse(CKD.stage...20 == '0.0', NA, CKD.stage...20)) %>%
  mutate(CKD.stage...20 = as.numeric(CKD.stage...20))
```

# Total Incidence

```{r}

fit <- survfit(Surv(age_pre, diagnosis) ~ 1, data = HNF1B_simple)

fit %>% 
  ggsurvfit(type='risk') +
  labs(
    x = 'Age at diagnosis',
    y = "Cumulative Incidence",
    title = "HNF1β nephropathy cumulative incidence plot."
  )+
  add_confidence_interval()+
  add_risktable()

ggsave('Survival_all.png', width = 10, height = 5)
```

0.5 incidence at 11 years. 50% atvejų nustatoma ties 11 metų.

# 5th CKD stage probability

```{r}
HNF1B_simple <- HNF1B_simple %>%
  mutate(severe_post = ifelse(CKD.stage...20 == 5, 1, 0)) %>%
  mutate(severe_pre = ifelse(CKD.stage...7 == 5, 1, 0))


fit <- survfit(Surv(age_post, severe_post) ~ 1, data = HNF1B_simple)

fit %>% 
  ggsurvfit() +
  labs(
    x = "Age at last follow-up",
    y = "Probability of not developing\n5th stage CKD",
    title = "Probability of not developing 5th stage CKD\nduring HNF1β nephropathy follow-up period."
  )+
  add_confidence_interval()+
  add_risktable()

ggsave('not_5th_stage.png', width = 10, height = 5)
```

# General mutation types Incidence

```{r}

HNF1B_survival <- HNF1B_simple %>%
  filter(!(mutacijos_tipas %in% c(NA, 'insertion', 'heterozygous intragenic mutation', 'frameshift', 'duplication', '?deletion'))) %>%
  filter(!is.na(mutacijos_tipas))

HNF1B_survival %>%
  group_by(mutacijos_tipas) %>%
  summarise(
    n=n()
  )

fit_mult <- survfit2(Surv(age_pre, diagnosis) ~ mutacijos_tipas, data = HNF1B_survival)

fit_mult %>% 
  ggsurvfit(type='risk', lwd=1) +
  labs(
    x = "Age at diagnosis",
    y = "Cumulative incidence",
    title = "Cumulative incidence of HNF1β nephropathy mutation types."
  )+
  theme(legend.position = "right")+
  scale_color_jco(labels = c("c_deletion, n=51", "missense, n=85", "nonsense, n=43", 'p_deletion, n=60', 'splice site, n=30', 'w_deletion, n=130'))

ggsave('Survival_mutations.png', width = 10, height = 5)
  
```

# GFG scatterplot

## eGFR \~Amžius

```{r}

extract_element_end <- function(string) {
  split_string <- strsplit(string, "[\\,]")
  element <- split_string[[1]][1]
  return(element)
}

HNF1B_simple$eGFR_pre <-as.numeric(lapply(HNF1B_simple$eGFR..mL.min.1.73m2....6, extract_element_end))

HNF1B_simple$eGFR_post <- as.numeric(lapply(HNF1B_simple$eGFR..mL.min.1.73m2....19, extract_element_end))

HNF1B_simple_prelm <- HNF1B_simple %>%
  filter(!is.na(eGFR_pre)) %>%
  filter(!is.na(age_pre)) 

r <- round(cor(HNF1B_simple_prelm$age_pre,HNF1B_simple_prelm$eGFR_pre), 2)
p <- cor.test(HNF1B_simple_prelm$age_pre,HNF1B_simple_prelm$eGFR_pre)$p.value

HNF1B_simple_prelm %>%
  dim()

HNF1B_simple_prelm %>%
  filter(!is.na(eGFR..mL.min.1.73m2....6)) %>%
  select(age_pre, eGFR_pre) %>%
  ggplot(aes(x=age_pre, y=eGFR_pre))+
  geom_point()+
  labs(
    y = 'eGFR, ml/min/1.73m\u00B2.',
    x = 'Age at diagnosis',
    title = "Scatter plot between age at HNF1β nephropathy diagnosis and eGFR (n = 294)."
  )+
  geom_smooth(method = 'lm', formula = y~x, se=FALSE)+
  annotate("text", x=40, y=150, label=paste0("Pearson's R = ", r), hjust=0)+
  annotate("text", x=40, y=135, label="P < 0.001", hjust=0)

ggsave('scatter_all.png', width = 10, height = 5)
```

```{r}
lm_fit <- lm(eGFR_pre ~ age_pre, data=HNF1B_simple)
summary(lm_fit)
```

Visgi eGFR daugiausiai priklauso nuo CKD stadijos

```{r}
lm_fit <- lm(eGFR_pre ~ age_pre + CKD.stage...7, data=HNF1B_simple)
summary(lm_fit)
```

```{r}
lm_fit <- lm(eGFR_pre ~ CKD.stage...7, data=HNF1B_simple)
summary(lm_fit)
```

## eGFR followup \~ Amžiaus skirtumas

```{r}
HNF1B_simple_difflm <- HNF1B_simple %>%
  filter(!is.na(eGFR_post)) %>%
  filter(!is.na(age_pre)) %>%
  filter(!is.na(age_post)) %>%
  mutate(age_diff = age_post - age_pre)

HNF1B_simple_difflm %>%
  dim()

r <- round(cor(HNF1B_simple_difflm$age_diff,HNF1B_simple_difflm$eGFR_post), 2)
p <- cor.test(HNF1B_simple_difflm$age_diff,HNF1B_simple_difflm$eGFR_post)$p.value
HNF1B_simple_difflm %>%
  filter(!is.na(eGFR_post)) %>%
  filter(!is.na(age_pre)) %>%
  filter(!is.na(age_post)) %>%
  select(age_diff, eGFR_post) %>%
  ggplot(aes(x=age_diff, y=eGFR_post))+
  geom_point()+
  labs(
    x = 'years between diagnosis and last follow-up',
    y = 'eGFR, ml/min/1.73m\u00B2.',
    title = "Scatter plot of eGFR and years between HNF1β nephropathy diagnosis and last follow-up (n = 97).",
  )+
  geom_smooth(method = 'lm', formula = y~x, se=FALSE)+
  annotate("text", x=15, y=100, label=paste0("Pearson's R = ", r), hjust=0) +
  annotate("text", x=15, y=90, label="P = 0.001", hjust=0)

ggsave('scatter_change.png', width = 10, height = 5)
```

# HNF1B mutacijos tipo įtaka CKD stadijai

Visos mutacijos

```{r}
HNF1B_simple %>%
  group_by(mutacijos_tipas) %>%
  summarise(
    n = n()
  )
```

Atrinktos mutacijos

```{r}
HNF1B_simple %>%
  select(mutacijos_tipas, CKD.stage...7) %>%
  filter(!(mutacijos_tipas %in% c(NA, 'insertion', 'heterozygous intragenic mutation', 'frameshift', 'duplication'))) %>%
  na.omit() %>%
  group_by(mutacijos_tipas) %>%
  summarise(
    n=n(),
    mean = mean(CKD.stage...7),
    sd = sd(CKD.stage...7)
  )

```

```{r output = FALSE}
bxp <- HNF1B_simple %>%
  filter(!(mutacijos_tipas %in% c(NA, 'insertion', 'heterozygous intragenic mutation', 'frameshift', 'duplication'))) %>%
  select(mutacijos_tipas, CKD.stage...7) %>%
  group_by(mutacijos_tipas) %>%
  ggboxplot(x='mutacijos_tipas', y='CKD.stage...7', fill='mutacijos_tipas')
bxp
```

```{r}
HNF1B_simple %>%
  select(mutacijos_tipas, CKD.stage...7) %>%
  filter(!(mutacijos_tipas %in% c(NA, 'insertion', 'heterozygous intragenic mutation', 'frameshift', 'duplication', '?deletion'))) %>%
  na.omit() %>%
  select(mutacijos_tipas, CKD.stage...7) %>%
  group_by(mutacijos_tipas) %>%
  summarise(
    mean = mean(CKD.stage...7),
    sd = sd(CKD.stage...7)
  ) %>%
  ggplot(aes(x=mutacijos_tipas, y=mean, fill=mutacijos_tipas))+
  geom_bar(stat = 'identity')+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                 position=position_dodge(.9))+
  labs(
    x = NULL,
    y = "CKD Stage",
    title = "Different CKD stage of patients according to the causative mutation type at diagnosis time of HNF1β nephropathy.",
    fill = "Mutation type"
  )+
  scale_fill_jco(labels = c("c_deletion, n=38", "missense, n=58", "nonsense, n=37", 'p_deletion, n=48', 'splice site, n=25', 'w_deletion, n=62'))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

ggsave('bar_mutations.png', width = 10, height = 5)
  

```

## ANOVA testas

```{r}

anova_data <- HNF1B_simple %>%
  select(mutacijos_tipas, CKD.stage...7) %>%
  filter(!(mutacijos_tipas %in% c(NA, 'insertion', 'heterozygous intragenic mutation', 'frameshift', 'duplication', '?deletion'))) %>%
  na.omit() %>%
  select(mutacijos_tipas, CKD.stage...7)

aov.m <- aov(CKD.stage...7 ~ mutacijos_tipas, data=anova_data)
summary(aov.m)
```

## Tukey post-hoc (pairwise comparisons)

```{r}
TukeyHSD(aov.m)
```

# CKD stage

## CKD diagnozės metu duomenys

```{r}

HNF1B_simple %>%
  filter(!(CKD.stage...7 == 0)) %>%
  filter(!is.na(CKD.stage...7)) %>%
  group_by(CKD.stage...7) %>%
  summarise(
    n=n()
  )


```

## Pacientai su CKD duomenim diagnozes ir follow-up metu

### diagnozes

```{r}
HNF1B_simple %>%
  filter(!(CKD.stage...7 == 0)) %>%
  filter(!is.na(CKD.stage...7)) %>%
  filter(!is.na(CKD.stage...20)) %>%
  group_by(CKD.stage...7) %>%
  summarise(
    n=n()
  )
```

### follow-up

```{r}
HNF1B_simple %>%
  filter(!(CKD.stage...7 == 0)) %>%
  filter(!is.na(CKD.stage...7)) %>%
  filter(!is.na(CKD.stage...20)) %>%
  group_by(CKD.stage...20) %>%
  summarise(
    n=n()
  )
```

### Pie charts

```{r}

set.seed(1)

pie.chart.data <- HNF1B_simple %>%
  filter(!(CKD.stage...7 == 0)) %>%
  filter(!is.na(CKD.stage...7)) %>%
  filter(!is.na(CKD.stage...20)) %>%
  mutate(CKD_pre = as.factor(CKD.stage...7)) %>%
  mutate(CKD_post = as.factor(CKD.stage...20)) %>%
  mutate(mutacijos_tipas = as.factor(mutacijos_tipas)) %>%
  select(Patient.ID, CKD_pre, CKD_post) %>%
  pivot_longer(c(CKD_pre, CKD_post), names_to = 'CKD_time', values_to = 'CKD_stage') %>%
  mutate(CKD_time= as.factor(CKD_time))
  # group_by(CKD_stage, CKD_time) %>%
  # summarise(
  #   n=n()
  # ) %>%

g1 <- ggpiestats(data = filter(pie.chart.data, CKD_time == 'CKD_pre'),
                 x=CKD_stage,
                 results.subtitle = FALSE,
                 label.repel = TRUE,
                 title='A. CKD at diagnosis',
                 label='both',
                 legend.title = "CKD Stage") + 
  scale_fill_manual(values=pal_jco("default")(10))
g2 <- ggpiestats(data = filter(pie.chart.data, CKD_time == 'CKD_post'),
                 x=CKD_stage,
                 results.subtitle = FALSE,
                 label.repel = TRUE,
                 title='B. CKD at last follow-up',
                 label='both',
                 legend.title = "CKD Stage")+ 
  scale_fill_manual(values=pal_jco("default")(10))


cqt <- CochranQTest(CKD_stage ~ CKD_time| Patient.ID, data=pie.chart.data)

cqt

plot1 <- ggarrange(g1, g2, ncol=2, nrow=1, common.legend = TRUE, legend='right')

plot1 <- set_palette(plot1, pal_jco("default")(10))

plot2 <- annotate_figure(plot1, bottom = text_grob("Cochran's Q test: X\u00B2(1)=15.114, p < 0.001; n = 68", hjust=0, vjust=-0.7))

plot3 <- annotate_figure(plot2, top = text_grob("nephropathy diagnosis and last follow-up (n = 68).", face='bold', hjust = 0.918, vjust=0.2, size = 16))

plot3 <- annotate_figure(plot3, top = text_grob("Changes in proportions of patients with different CKD stages at HNF1β", face='bold', hjust = 0.65, vjust=0.2, size = 16))

plot3

ggsave('pie_charts.png', plot=plot3, height = 5, width = 10, bg = 'white')
```

### Histogram

Žiūrime kaip keičiasi CKD lygis nuo diagnosis iki follow-up

```{r}
fal = -3:3

HNF1B_simple %>%
  filter(!(CKD.stage...7 == 0)) %>%
  filter(!is.na(CKD.stage...7)) %>%
  filter(!is.na(CKD.stage...20)) %>%
  mutate(CKD_diff = CKD.stage...20-CKD.stage...7) %>%
  mutate(mutacijos_tipas = as.factor(mutacijos_tipas)) %>%
  select(Patient.ID, CKD_diff) %>%
  ggplot(aes(CKD_diff))+
  geom_bar()+
  scale_x_continuous('Change in CKD stage', breaks= fal)+
  labs(
    y = 'Patient count',
    title = 'CKD severity change from HNF1β nephropathy diagnosis to last follow-up.'
  )

ggsave('hist_change.png', width = 10, height = 5)

```

# Categorical Data

## Diabetes

```{r}
HNF1B_simple <- HNF1B_simple %>%
  mutate(
  Proteinuria..yes.no..g.l....9 = ifelse(grepl('yes', Proteinuria..yes.no..g.l....9, ignore.case = TRUE), 'yes', 'no'),
  Diabetes..yes.no..type....10 = ifelse(grepl('yes', Diabetes..yes.no..type....10, ignore.case = TRUE), 'yes', 'no'),
  Hyperuricemia..yes.no..mg.dL....11 = ifelse(grepl('yes', Hyperuricemia..yes.no..mg.dL....11, ignore.case = TRUE), 'yes', 'no'),
  Hypomagnesemia..yes.NA..mmol.l. = ifelse(grepl('yes', Hypomagnesemia..yes.NA..mmol.l., ignore.case = TRUE), 'yes', 'no'),
  HTN..yes.no....8 = ifelse(grepl('yes', HTN..yes.no....8, ignore.case = TRUE), 'yes', 'no')
         )
```

```{r}
library(ggstatsplot)
proteinuria <- HNF1B_simple %>%
  select(mutacijos_tipas, Proteinuria..yes.no..g.l....9, Hyperuricemia..yes.no..mg.dL....11, Diabetes..yes.no..type....10, HTN..yes.no....8,Hypomagnesemia..yes.NA..mmol.l.) %>%
  filter(mutacijos_tipas %in% c('cdeletion', 'wdeletion','pdeletion','missense','nonsense', 'splice')) %>%
  na.omit() %>%
  mutate(mutacijos_tipas = as.factor(mutacijos_tipas)) %>%
  mutate(Proteinuria = as.factor(Proteinuria..yes.no..g.l....9)) %>%
  mutate(Hyperuricemia = as.factor(Hyperuricemia..yes.no..mg.dL....11)) %>%
  mutate(Diabetes = as.factor(Diabetes..yes.no..type....10)) %>%
  mutate(HTN = as.factor(HTN..yes.no....8)) %>%
  mutate(Hypomagnesemia = as.factor(Hypomagnesemia..yes.NA..mmol.l.))

g1 <- ggstatsplot::ggbarstats(
  data = proteinuria,
  x=Diabetes,
  y=mutacijos_tipas,
  proportion.test = NULL,
  package      = "wesanderson",
  palette      = "Darjeeling2",
  ggtheme = theme_pubclean(),
  bf.message=FALSE,
  messages=FALSE,
  xlab = "",
  type='parametric'
  )

g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL

g1 + theme(legend.position = 'right')

ggsave('diabetes.png', width = 6, height=4)

```

## Proteinuria

```{r}
g1 <- ggstatsplot::ggbarstats(
  data = proteinuria,
  x=Proteinuria,
  y=mutacijos_tipas,
  proportion.test = NULL,
  package      = "wesanderson",
  palette      = "Darjeeling2",
  ggtheme = theme_pubclean(),
  bf.message=FALSE,
  messages=FALSE,
  xlab = "",
  type='parametric'
  )

g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL

g1 + theme(legend.position = 'right')

ggsave('proteinuria.png', width = 6, height=4)
```

## Hyperuricemia

```{r}
g1 <- ggstatsplot::ggbarstats(
  data = proteinuria,
  x=Hyperuricemia,
  y=mutacijos_tipas,
  proportion.test = NULL,
  package      = "wesanderson",
  palette      = "Darjeeling2",
  ggtheme = theme_pubclean(),
  bf.message=FALSE,
  messages=FALSE,
  xlab = "",
  type='parametric'
  )

g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL

g1 + theme(legend.position = 'right')

ggsave('hyperuricemia.png', width = 6, height=4)
```

## HTN

```{r}
g1 <- ggstatsplot::ggbarstats(
  data = proteinuria,
  x=HTN,
  y=mutacijos_tipas,
  proportion.test = NULL,
  package      = "wesanderson",
  palette      = "Darjeeling2",
  ggtheme = theme_pubclean(),
  bf.message=FALSE,
  messages=FALSE,
  xlab = "",
  type='parametric'
  )

g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL

g1 + theme(legend.position = 'right')

ggsave('HTN.png', width = 6, height=4)
```

## Hypomagnesemia

```{r}
g1 <- ggstatsplot::ggbarstats(
  data = proteinuria,
  x=Hypomagnesemia,
  y=mutacijos_tipas,
  proportion.test = NULL,
  package      = "wesanderson",
  palette      = "Darjeeling2",
  ggtheme = theme_pubclean(),
  bf.message=FALSE,
  messages=FALSE,
  xlab = "",
  type='parametric'
  )

g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL
g1$labels$subtitle[[4]] <- NULL

g1 + theme(legend.position = 'right')

ggsave('hypomagnesemia.png', width = 6, height=4)
```

## Age \~ Mutation Type

```{r}
age_mut_plot <- HNF1B_simple %>%
  select(mutacijos_tipas, age_pre) %>%
  na.omit() %>%
  filter(mutacijos_tipas %in% c('cdeletion', 'wdeletion','pdeletion','missense','nonsense', 'splice')) %>%
  group_by(mutacijos_tipas) %>%
  ggbetweenstats(data=., x = mutacijos_tipas, y=age_pre, 
                 p.adjust.method = 'bonferroni', 
                 bf.message = FALSE, 
                 ggtheme = theme_pubr(), 
                 palette = 'default_jco', 
                 package='ggsci', 
                 type = 'parametric', 
                 var.equal = TRUE, 
                 xlab = "", 
                 ylab = 'Age at diagnosis',
                 # ggplot.component = list(ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis(name = "Pairwise test: Student's t")))
  )

age_mut_plot$labels$subtitle[[4]] <- NULL
age_mut_plot$labels$subtitle[[4]] <- NULL
age_mut_plot$labels$subtitle[[4]] <- NULL

age_mut_plot

ggsave('age_mutation.png', width = 8, height = 7)
```
